### 2014-05-14 ApTranscriptome_TR.Rmd
Updated T~on~ analysis.
Merged both *Low* and *High* onto same plot - better visualization of when genes turned on.

**May 6, 2014**

@@

# Load libraries
library(R.utils)
library(ggplot2)
library(knitr)
library(knitcitations)
library(pander)
library(stringr)
library(data.table)
library(RCurl)
library(plyr)
library(MASS)
library(xtable)
library(methods)

@@

    sailfish -i sailfish-index-Trinity-cap3-uclust -o sailfish-expression-Trinity-cap3-uclust/A22-0 -l "T=PE:O=><:S=U" -r A22-0_ATCACG.unpaired.left.fastq A22-0_ATCACG.unpaired.right.fastq -1 A22-0_ATCACG.paired.left.fastq -2 A22-0_ATCACG.paired.right.fastq  -p 4

@@

```{r RxNlmAIC, echo=TRUE, eval=TRUE, results='hide'}
# perform model selection for responsive transcripts
RxNlmAIC <- try(dlply(sig.TPM.dt.sub, .(Transcript), lmFunc))
```

@@

# save results to file
write.table(file = paste(resultsdir, "Ap_responsive_transcripts.txt", sep = ""), Ap.response.type, row.names = FALSE, col.names = TRUE, quote = FALSE, sep = "\t")
```

@@

```{r optimum_expression_comparison, cache=TRUE, eval=TRUE}
# list of transcripts that are 'high' expressed in A22
A22.high.transcripts <- Ap.response.type[which(Ap.response.type$A22.type == "High"), ]

@@

```{r A22_high_wilcoxon}
(w1 <- wilcox.test(A22.high.transcripts$A22.opt, A22.high.transcripts$Ar.opt, alternative = "two.sided", paired = TRUE, conf.int = TRUE))
```

@@

# compare expression at optimum temp (Ar.opt) between colonies using t-test
t.test(Ar.low.transcripts$A22.opt, Ar.low.transcripts$Ar.opt)
boxplot(Ar.low.transcripts$A22.opt, Ar.low.transcripts$Ar.opt)
# t-test with log values
t.test(log(Ar.low.transcripts$A22.opt+1), log(Ar.low.transcripts$Ar.opt+1))
boxplot(log(Ar.low.transcripts$A22.opt+1), log(Ar.low.transcripts$Ar.opt+1))

# Wilcoxon signed rank-test
(w2 <- wilcox.test(Ar.low.transcripts$A22.opt, Ar.low.transcripts$Ar.opt, alternative = "two.sided", paired = TRUE, conf.int = TRUE))
```

@@

```{r intermediate_transcripts}
# list of transcripts that are 'Intermediate' expressed in Ar
Ar.int.transcripts <- Ap.response.type[which(Ap.response.type$Ar.type == "Intermediate"), ]
# T test
t.test(Ar.int.transcripts$A22.opt, Ar.int.transcripts$Ar.opt)
t.test(log(Ar.int.transcripts$A22.opt+1), log(Ar.int.transcripts$Ar.opt+1))
# Wilcoxon signed rank-test
wilcox.test(Ar.int.transcripts$A22.opt, Ar.int.transcripts$Ar.opt, alternative = "two.sided", paired = TRUE, conf.int = TRUE)

# list of transcripts that are 'Intermediate' expressed in A22
A22.int.transcripts <- Ap.response.type[which(Ap.response.type$A22.type == "Intermediate"), ]
# T test
t.test(A22.int.transcripts$A22.opt, A22.int.transcripts$Ar.opt)
t.test(log(A22.int.transcripts$A22.opt+1), log(A22.int.transcripts$Ar.opt+1))
# Wilcoxon signed rank-test
wilcox.test(A22.int.transcripts$A22.opt, A22.int.transcripts$Ar.opt, alternative = "two.sided", paired = TRUE, conf.int = TRUE)
```

@@

Predict for responsive transcripts

```{r responsive_pred}

predFunc <- function(foo) {
  # Function supplied to `ddply` to predict expression at each temperature
  #
  # Args:
  #  foo: data.frame model is fit to
  #
  #  'model' must be specified in environment  
  #
  # Returns:
  #  data.frame of Transcripts with pval

  lmout <- eval(parse(text = paste("lm(", model, ", data = foo)", sep = "")))

  # predict values
  pout <- predict(lmout)
  # back-transform from log-scale
  foo$pTPM <- exp(pout)

  # prediction can give values very close to zero or negative. biologically, these are meaningless so change to zero
  foo$pTPM <- ifelse(foo$pTPM < 0, 0, round(foo$pTPM, 5))
  # plot
  #with(foo, qplot(val, TPM, colour = colony, geom = "smooth"))
  #with(foo, qplot(val, pTPM, colour = colony, geom = "smooth"))

  pred.out.df <- data.frame(Transcript = foo$Transcript, TPM = foo$TPM, val = foo$val, colony = foo$colony, pTPM = foo$pTPM)
}

# apply predFunc to all responsive transcripts
resp.TPM.dt.sub.pred <- ddply(resp.TPM.dt.sub, .(Transcript), .inform="TRUE", predFunc)

# setkey to Transcript and colony 
resp.TPM.dt.sub.pred <- data.table(resp.TPM.dt.sub.pred)
setkey(resp.TPM.dt.sub.pred, Transcript, colony)

# transcripts expressed at *High* temperatures in A22
A22.high.TPM.dt.sub <- resp.TPM.dt.sub.pred[J(A22.high, "A22")]
str(A22.high.TPM.dt.sub)

# make data.frame for results
A22.high.T_on <- data.frame(Transcript = unique(A22.high.TPM.dt.sub$Transcript), colony = rep(A22.high.TPM.dt.sub$colony, length = length(unique(A22.high.TPM.dt.sub$Transcript))), T_on = NA, pT_on = NA)

@@

for(i in unique(A22.high.TPM.dt.sub$Transcript)) {
    subdf <- A22.high.TPM.dt.sub[i]
    subdf <- subdf[which(subdf$val > 14), ]
    T_on <- subdf[median(which(diff(subdf$TPM) == max(diff(subdf$TPM))))+1, val]
    pT_on <- subdf[median(which(diff(subdf$pTPM) == max(diff(subdf$pTPM))))+1, val]
    A22.high.T_on[which(A22.high.T_on$Transcript == i), "T_on"] <- T_on
    A22.high.T_on[which(A22.high.T_on$Transcript == i), "pT_on"] <- pT_on
}

# repeat for Ar
Ar.high.TPM.dt.sub <- resp.TPM.dt.sub.pred[J(Ar.high, "Ar")]
str(Ar.high.TPM.dt.sub)

Ar.high.T_on <- data.frame(Transcript = unique(Ar.high.TPM.dt.sub$Transcript), colony = rep(Ar.high.TPM.dt.sub$colony, length = length(unique(Ar.high.TPM.dt.sub$Transcript))), T_on = NA, pT_on = NA)

for(i in unique(Ar.high.TPM.dt.sub$Transcript)) {
    subdf <- Ar.high.TPM.dt.sub[i]
    subdf <- subdf[which(subdf$val > 14), ]
    T_on <- subdf[median(which(diff(subdf$TPM) == max(diff(subdf$TPM))))+1, val]
    pT_on <- subdf[median(which(diff(subdf$pTPM) == max(diff(subdf$pTPM))))+1, val]
    Ar.high.T_on[which(Ar.high.T_on$Transcript == i), "T_on"] <- T_on
    Ar.high.T_on[which(Ar.high.T_on$Transcript == i), "pT_on"] <- pT_on
}

cor.test(Ar.high.T_on$T_on, Ar.high.T_on$pT_on)
cor.test(A22.high.T_on$T_on, A22.high.T_on$pT_on)

t.test(Ar.high.T_on$T_on, A22.high.T_on$T_on)
t.test(Ar.high.T_on$pT_on, A22.high.T_on$pT_on)
```

Genes with increased expression at *High* temperatures are turned on at higher temperatures in *ApVT* than *AcNC*.

```{r plot_high_T_on, echo=FALSE, eval=TRUE}
# prep data for plotting
Ap.high.T_on <- rbind(A22.high.T_on, Ar.high.T_on)

# plot
png("T_on_high.png")
T_on_plot <- ggplot(Ap.high.T_on, aes(x=T_on, fill=colony)) + geom_density(alpha=0.2, position="identity")
T_on_plot + scale_y_continuous(name="Density") +
  scale_x_continuous(name=expression("Temperature"))
dev.off()

# plot predicted
png("pT_on_high.png")
pT_on_plot <- ggplot(Ap.high.T_on, aes(x=pT_on, fill=colony)) + geom_density(alpha=0.2, position="identity")
pT_on_plot + scale_y_continuous(name="Density") +
  scale_x_continuous(name=expression("Temperature"))
dev.off()
```

@@

```{r echo=FALSE, results='hide'}
responsive.lms.ann.type <- signif.transcripts[names(responsive.lms)]
# combine responsive.lms.ann with "type" information
responsive.lms.ann.type <- merge(responsive.lms.ann.type, Ap.response.type[,c("Transcript", "Ar.type", "Ar.type")], by = "Transcript", all = TRUE)
str(responsive.lms.ann.type)
write.csv(responsive.lms.ann.type, file = "results/Ap_responsive_genes_20140409.csv", row.names = FALSE)
```

@@

```{r gsea_all}
# create geneList. note that NA values cause problems with topGO
# so set any NA to 1 as need to retain for GO analysis
Ap.geneList <- RxNpval$padj
Ap.geneList[which(is.na(Ap.geneList))] <- 1
stopifnot(length(which(is.na(Ap.geneList))) == 0)
names(Ap.geneList) <- RxNpval$Transcript
str(Ap.geneList)

@@

```{r gsea_table, echo=FALSE, eval=TRUE, results='asis'}
## merge results into single table
Ap.high.gsea$Type <- "High"
Ap.low.gsea$Type <- "Low"
Ap.bim.gsea$Type <- "Bimodal"
Ap.int.gsea$Type <- "Intermediate"
# combine
Ap.gsea <- rbind(Ap.high.gsea, Ap.low.gsea, Ap.bim.gsea, Ap.int.gsea)
# reorder
Ap.gsea1 <- Ap.gsea[,c("Type", "GO.ID", "Term", "Annotated", "Significant", "Expected", "parentchild")]
colnames(Ap.gsea1)[7] <- "P"
# display
Ap.gsea1
```

@@

```{r gsea_high_A22}
A22.high <- Ap.response.type[which(Ap.response.type$A22.type == "High" & Ap.response.type$Ar.type != "High"), "Transcript"]
# create gene list, setting value to 1 for "bim" transcripts
A22.geneList.high <- rep(0, length = length(Ap.geneList))
names(A22.geneList.high) <- names(Ap.geneList)
A22.geneList.high[(which(names(A22.geneList.high) %in% A22.high))] <- 1
# check correct number of values set to 1
table(A22.geneList.high)

@@

```{r gsea_high_Ar}
Ar.high <- Ap.response.type[which(Ap.response.type$A22.type != "High" & Ap.response.type$Ar.type == "High"), "Transcript"]
# create gene list, setting value to 1 for "bim" transcripts
Ar.geneList.high <- rep(0, length = length(Ap.geneList))
names(Ar.geneList.high) <- names(Ap.geneList)
Ar.geneList.high[(which(names(Ar.geneList.high) %in% Ar.high))] <- 1
# check correct number of values set to 1
table(Ar.geneList.high)

@@

```{r gsea_low_A22}
A22.low <- Ap.response.type[which(Ap.response.type$A22.type == "Low" & Ap.response.type$Ar.type != "Low"), "Transcript"]
# create gene list, setting value to 1 for "bim" transcripts
A22.geneList.low <- rep(0, length = length(Ap.geneList))
names(A22.geneList.low) <- names(Ap.geneList)
A22.geneList.low[(which(names(A22.geneList.low) %in% A22.low))] <- 1
# check correct number of values set to 1
table(A22.geneList.low)

@@

```{r gsea_low_Ar}
Ar.low <- Ap.response.type[which(Ap.response.type$A22.type != "Low" & Ap.response.type$Ar.type == "Low"), "Transcript"]
# create gene list, setting value to 1 for "bim" transcripts
Ar.geneList.low <- rep(0, length = length(Ap.geneList))
names(Ar.geneList.low) <- names(Ap.geneList)
Ar.geneList.low[(which(names(Ar.geneList.low) %in% Ar.low))] <- 1
# check correct number of values set to 1
table(Ar.geneList.low)

@@

```{r gsea_bim_A22}
A22.bim <- Ap.response.type[which(Ap.response.type$A22.type == "Bimodal" & Ap.response.type$Ar.type != "Bimodal"), "Transcript"]
# create gene list, setting value to 1 for "bim" transcripts
A22.geneList.bim <- rep(0, length = length(Ap.geneList))
names(A22.geneList.bim) <- names(Ap.geneList)
A22.geneList.bim[(which(names(A22.geneList.bim) %in% A22.bim))] <- 1
# check correct number of values set to 1
table(A22.geneList.bim)

@@

```{r gsea_bim_Ar}
Ar.bim <- Ap.response.type[which(Ap.response.type$A22.type != "Bimodal" & Ap.response.type$Ar.type == "Bimodal"), "Transcript"]
# create gene list, setting value to 1 for "bim" transcripts
Ar.geneList.bim <- rep(0, length = length(Ap.geneList))
names(Ar.geneList.bim) <- names(Ap.geneList)
Ar.geneList.bim[(which(names(Ar.geneList.bim) %in% Ar.bim))] <- 1
# check correct number of values set to 1
table(Ar.geneList.bim)

@@

```{r gsea_int_A22}
A22.int <- Ap.response.type[which(Ap.response.type$A22.type == "Intermediate" & Ap.response.type$Ar.type != "Intermediate"), "Transcript"]
# create gene list, setting value to 1 for "bim" transcripts
A22.geneList.int <- rep(0, length = length(Ap.geneList))
names(A22.geneList.int) <- names(Ap.geneList)
A22.geneList.int[(which(names(A22.geneList.int) %in% A22.int))] <- 1
# check correct number of values set to 1
table(A22.geneList.int)

@@

```{r gsea_int_Ar}
Ar.int <- Ap.response.type[which(Ap.response.type$A22.type != "Intermediate" & Ap.response.type$Ar.type == "Intermediate"), "Transcript"]
# create gene list, setting value to 1 for "bim" transcripts
Ar.geneList.int <- rep(0, length = length(Ap.geneList))
names(Ar.geneList.int) <- names(Ap.geneList)
Ar.geneList.int[(which(names(Ar.geneList.int) %in% Ar.int))] <- 1
# check correct number of values set to 1
table(Ar.geneList.int)

@@

```{r rData, echo=FALSE}
# save RData for manuscript file
save(tt2, Ap.df, w1, w2, file = "ms.RData")
```

>>>>

**May 14, 2014**

@@

# Load libraries
library(R.utils)
library(ggplot2)
library(knitr)
library(knitcitations)
library(pander)
library(stringr)
library(data.table)
library(RCurl)
library(plyr)
library(MASS)
library(xtable)
library(methods) # required by `knitcitations` when file is `knit` file from command-line

@@

    sailfish quant -i sailfish-index-Trinity-cap3-uclust -o sailfish-expression-Trinity-cap3-uclust/A22-0 -l "T=SE:S=U" -r A22-0_ATCACG.unpaired.left.fastq A22-0_ATCACG.unpaired.right.fastq A22-0_ATCACG.paired.left.fastq A22-0_ATCACG.paired.right.fastq -p 4

While it is possible to separately specify the paired-end and orphaned single-end reads in Sailfish v0.6.3, the results are exactly the same as if they are all entered as SE.

@@

```{r RxNlmAIC, echo=TRUE, eval=TRUE, results='hide'}
# perform model selection for responsive transcripts
# need to use `try` to avoid stopping on error for AIC at Infinity
RxNlmAIC <- try(dlply(sig.TPM.dt.sub, .(Transcript), lmFunc))
```

@@

# save results to file
write.table(file = paste(resultsdir, "Ap-responsive-transcripts", Sys.Date(), ".txt", sep = ""), Ap.response.type, row.names = FALSE, col.names = TRUE, quote = FALSE, sep = "\t")
```

@@

```{r optimum_expression_comparison}
# list of transcripts that are 'high' expressed in A22
A22.high.transcripts <- Ap.response.type[which(Ap.response.type$A22.type == "High"), ]

@@

```{r A22_high_wilcoxon}
w1 <- wilcox.test(A22.high.transcripts$A22.opt, A22.high.transcripts$Ar.opt, alternative = "two.sided", paired = TRUE, conf.int = TRUE)
w1
```

@@

# t-test with log values
t.test(log(Ar.low.transcripts$A22.opt+1), log(Ar.low.transcripts$Ar.opt+1))
boxplot(log(Ar.low.transcripts$A22.opt+1), log(Ar.low.transcripts$Ar.opt+1))

# Wilcoxon signed rank-test
w2 <- wilcox.test(Ar.low.transcripts$A22.opt, Ar.low.transcripts$Ar.opt, alternative = "two.sided", paired = TRUE, conf.int = TRUE)
w2
```

@@

```{r intermediate_transcripts}
# list of transcripts that are 'Intermediate' expressed in Ar
Ar.int.transcripts <- Ap.response.type[which(Ap.response.type$Ar.type == "Intermediate"), ]
# T test
t.test(log(Ar.int.transcripts$A22.opt+1), log(Ar.int.transcripts$Ar.opt+1))
# Wilcoxon signed rank-test
wilcox.test(Ar.int.transcripts$A22.opt, Ar.int.transcripts$Ar.opt, alternative = "two.sided", paired = TRUE, conf.int = TRUE)

# list of transcripts that are 'Intermediate' expressed in A22
A22.int.transcripts <- Ap.response.type[which(Ap.response.type$A22.type == "Intermediate"), ]
# T test
t.test(log(A22.int.transcripts$A22.opt+1), log(A22.int.transcripts$Ar.opt+1))
# Wilcoxon signed rank-test
wilcox.test(A22.int.transcripts$A22.opt, A22.int.transcripts$Ar.opt, alternative = "two.sided", paired = TRUE, conf.int = TRUE)
```

@@

Predict expression for responsive transcripts.

```{r responsive_pred}
# apply predFunc to all responsive transcripts
resp.TPM.dt.sub.pred <- ddply(resp.TPM.dt.sub, .(Transcript), .inform="TRUE", predFunc)

# setkey to Transcript and colony 
resp.TPM.dt.sub.pred <- data.table(resp.TPM.dt.sub.pred)
setkey(resp.TPM.dt.sub.pred, Transcript, colony)
```

For next analyses, extract list of gene names by response type.

```{r extractionTranscripts}
A22.high <- Ap.response.type[which(Ap.response.type$A22.type == "High" & Ap.response.type$Ar.type != "High"), "Transcript"]
Ar.high <- Ap.response.type[which(Ap.response.type$A22.type != "High" & Ap.response.type$Ar.type == "High"), "Transcript"]

A22.low <- Ap.response.type[which(Ap.response.type$A22.type == "Low" & Ap.response.type$Ar.type != "Low"), "Transcript"]
Ar.low <- Ap.response.type[which(Ap.response.type$A22.type != "Low" & Ap.response.type$Ar.type == "Low"), "Transcript"]

A22.bim <- Ap.response.type[which(Ap.response.type$A22.type == "Bimodal" & Ap.response.type$Ar.type != "Bimodal"), "Transcript"]
Ar.bim <- Ap.response.type[which(Ap.response.type$A22.type != "Bimodal" & Ap.response.type$Ar.type == "Bimodal"), "Transcript"]

A22.int <- Ap.response.type[which(Ap.response.type$A22.type == "Intermediate" & Ap.response.type$Ar.type != "Intermediate"), "Transcript"]
Ar.int <- Ap.response.type[which(Ap.response.type$A22.type != "Intermediate" & Ap.response.type$Ar.type == "Intermediate"), "Transcript"]
```

Calculate T~on~ for *High* genes in each colony.

```{r T_on_high}
# transcripts expressed at *High* and *Bimodal* temperatures in A22
A22.high.TPM.dt.sub <- resp.TPM.dt.sub.pred[J(union(A22.high, A22.bim), "A22")]
setkey(A22.high.TPM.dt.sub, Transcript)
str(A22.high.TPM.dt.sub)

# make data.frame for results
l1 <- length(unique(A22.high.TPM.dt.sub$Transcript))
A22.high.T_on <- data.frame(Transcript = unique(A22.high.TPM.dt.sub$Transcript), colony = rep("ApVT", length = l1), type = rep("High", length = l1), T_on = NA)

@@

for(i in unique(A22.high.TPM.dt.sub$Transcript)) {
    subdf <- A22.high.TPM.dt.sub[i]
    subdf <- subdf[which(subdf$val > 14), ]
    T_on <- subdf[median(which(diff(subdf$TPM) == max(diff(subdf$TPM))))+1, val]
    A22.high.T_on[which(A22.high.T_on$Transcript == i), "T_on"] <- T_on
}

# repeat for Ar
Ar.high.TPM.dt.sub <- resp.TPM.dt.sub.pred[J(union(Ar.high, Ar.bim), "Ar")]
setkey(Ar.high.TPM.dt.sub, Transcript)
str(Ar.high.TPM.dt.sub)

l2 <- length(unique(Ar.high.TPM.dt.sub$Transcript))
Ar.high.T_on <- data.frame(Transcript = unique(Ar.high.TPM.dt.sub$Transcript), colony = rep("AcNC", length = l2), type = rep("High", length = l2), T_on = NA)

for(i in unique(Ar.high.TPM.dt.sub$Transcript)) {
    subdf <- Ar.high.TPM.dt.sub[i]
    subdf <- subdf[which(subdf$val > 14), ]
    T_on <- subdf[median(which(diff(subdf$TPM) == max(diff(subdf$TPM))))+1, val]
    Ar.high.T_on[which(Ar.high.T_on$Transcript == i), "T_on"] <- T_on
}   
```

Determine if T~on~ is greater in *A22* or *Ar* for *High* genes.

```{r T_on_high_ttest}
t.test(Ar.high.T_on$T_on, A22.high.T_on$T_on)
```

Genes with increased expression at *High* temperatures are turned on at, on average, 1°C higher in *AcNC* than *ApVT*.

Repeat analysis for *Low* genes.

```{r T_on_low}
# transcripts expressed at *Low* temperatures in A22
A22.low.TPM.dt.sub <- resp.TPM.dt.sub.pred[J(A22.low, "A22")]
str(A22.low.TPM.dt.sub)

intersect(A22.low.TPM.dt.sub, A22.high.TPM.dt.sub)
# Good - no overlap

# make data.frame for results
l3 <- length(unique(A22.low.TPM.dt.sub$Transcript))
A22.low.T_on <- data.frame(Transcript = unique(A22.low.TPM.dt.sub$Transcript), colony = rep("ApVT", length = l3), type = rep("Low", length = l3), T_on = NA)

for(i in unique(A22.low.TPM.dt.sub$Transcript)) {
    subdf <- A22.low.TPM.dt.sub[i]
    subdf <- subdf[which(subdf$val < 21), ]
    T_on <- subdf[median(which(diff(subdf$TPM) == max(diff(subdf$TPM))))+1, val]
    A22.low.T_on[which(A22.low.T_on$Transcript == i), "T_on"] <- T_on
}

# repeat for Ar
Ar.low.TPM.dt.sub <- resp.TPM.dt.sub.pred[J(Ar.low, "Ar")]
str(Ar.low.TPM.dt.sub)

intersect(A22.low.TPM.dt.sub, A22.high.TPM.dt.sub)
# Good - no overlap

l4 <- length(unique(Ar.low.TPM.dt.sub$Transcript))
Ar.low.T_on <- data.frame(Transcript = unique(Ar.low.TPM.dt.sub$Transcript), colony = rep("AcNC", length = l4), type = rep("Low", length = length(unique(Ar.low.TPM.dt.sub$Transcript))), T_on = NA)

for(i in unique(Ar.low.TPM.dt.sub$Transcript)) {
    subdf <- Ar.low.TPM.dt.sub[i]
    subdf <- subdf[which(subdf$val < 21), ]
    T_on <- subdf[median(which(diff(subdf$TPM) == max(diff(subdf$TPM))))+1, val]
    Ar.low.T_on[which(Ar.low.T_on$Transcript == i), "T_on"] <- T_on
}

t.test(Ar.low.T_on$T_on, A22.low.T_on$T_on)
```

Genes with increased expression at *Low* temperatures are turned on at lower temperatures in *AcNC* than *ApVT*.

Visualize T~on~ for both *Low* and *High* genes on the same plot.

```{r plot_low_T_on, echo=FALSE, eval=TRUE}
# merge data for plotting
Ap.T_on <- rbind(A22.high.T_on, Ar.high.T_on, A22.low.T_on, Ar.low.T_on)

# plot
T_on_plot <- ggplot(Ap.T_on, aes(x = T_on, y = ..density.., colour = colony, group = colony, fill = colony)) +
    geom_histogram(position = "dodge", binwidth = 3) +
    geom_density(adjust = 3, fill = NA, size = 2) + 
    scale_y_continuous(name="Density") +
    scale_x_continuous(name=expression("Temperature"))
print(T_on_plot)
```

@@

```{r echo=FALSE, results='hide'}
responsive.lms.ann.type <- signif.transcripts[names(responsive.lms)]
# combine responsive.lms.ann with "type" information
responsive.lms.ann.type <- merge(responsive.lms.ann.type, Ap.response.type[,c("Transcript", "Ar.type", "Ar.type")], by = "Transcript", all = TRUE)
str(responsive.lms.ann.type)
write.csv(responsive.lms.ann.type, file = paste(resultsdir, "Ap_responsive_genes_", Sys.Date(), ".csv", sep = ""), row.names = FALSE)
```

@@

```{r gsea_all, results='hide'}
# create geneList. note that NA values cause problems with topGO
# so set any NA to 1 as need to retain for GO analysis
Ap.geneList <- RxNpval$padj
Ap.geneList[which(is.na(Ap.geneList))] <- 1
stopifnot(length(which(is.na(Ap.geneList))) == 0)
names(Ap.geneList) <- RxNpval$Transcript
str(Ap.geneList)

@@

```{r gsea_table, echo=FALSE, eval=TRUE, results='asis'}
## merge results into single table
Ap.high.gsea$Type <- "High"
Ap.low.gsea$Type <- "Low"
Ap.bim.gsea$Type <- "Bimodal"
Ap.int.gsea$Type <- "Intermediate"
# combine
Ap.gsea <- rbind(Ap.high.gsea, Ap.low.gsea, Ap.bim.gsea, Ap.int.gsea)
# reorder
Ap.gsea1 <- Ap.gsea[,c("Type", "GO.ID", "Term", "Annotated", "Significant", "Expected", "parentchild")]
colnames(Ap.gsea1)[7] <- "P"
# display
Ap.gsea1[ , 1:3]
```

@@

```{r gsea_high_A22}
# create gene list, setting value to 1 for "bim" transcripts
A22.geneList.high <- rep(0, length = length(Ap.geneList))
names(A22.geneList.high) <- names(Ap.geneList)
A22.geneList.high[(which(names(A22.geneList.high) %in% A22.high))] <- 1
# check correct number of values set to 1
table(A22.geneList.high)

@@

```{r gsea_high_Ar}
# create gene list, setting value to 1 for "bim" transcripts
Ar.geneList.high <- rep(0, length = length(Ap.geneList))
names(Ar.geneList.high) <- names(Ap.geneList)
Ar.geneList.high[(which(names(Ar.geneList.high) %in% Ar.high))] <- 1
# check correct number of values set to 1
table(Ar.geneList.high)

@@

```{r gsea_low_A22}
# create gene list, setting value to 1 for "bim" transcripts
A22.geneList.low <- rep(0, length = length(Ap.geneList))
names(A22.geneList.low) <- names(Ap.geneList)
A22.geneList.low[(which(names(A22.geneList.low) %in% A22.low))] <- 1
# check correct number of values set to 1
table(A22.geneList.low)

@@

```{r gsea_low_Ar}
# create gene list, setting value to 1 for "bim" transcripts
Ar.geneList.low <- rep(0, length = length(Ap.geneList))
names(Ar.geneList.low) <- names(Ap.geneList)
Ar.geneList.low[(which(names(Ar.geneList.low) %in% Ar.low))] <- 1
# check correct number of values set to 1
table(Ar.geneList.low)

@@

```{r gsea_bim_A22}
# create gene list, setting value to 1 for "bim" transcripts
A22.geneList.bim <- rep(0, length = length(Ap.geneList))
names(A22.geneList.bim) <- names(Ap.geneList)
A22.geneList.bim[(which(names(A22.geneList.bim) %in% A22.bim))] <- 1
# check correct number of values set to 1
table(A22.geneList.bim)

@@

```{r gsea_bim_Ar}
# create gene list, setting value to 1 for "bim" transcripts
Ar.geneList.bim <- rep(0, length = length(Ap.geneList))
names(Ar.geneList.bim) <- names(Ap.geneList)
Ar.geneList.bim[(which(names(Ar.geneList.bim) %in% Ar.bim))] <- 1
# check correct number of values set to 1
table(Ar.geneList.bim)

@@

```{r gsea_int_A22}
# create gene list, setting value to 1 for "bim" transcripts
A22.geneList.int <- rep(0, length = length(Ap.geneList))
names(A22.geneList.int) <- names(Ap.geneList)
A22.geneList.int[(which(names(A22.geneList.int) %in% A22.int))] <- 1
# check correct number of values set to 1
table(A22.geneList.int)

@@

```{r gsea_int_Ar}
# create gene list, setting value to 1 for "bim" transcripts
Ar.geneList.int <- rep(0, length = length(Ap.geneList))
names(Ar.geneList.int) <- names(Ap.geneList)
Ar.geneList.int[(which(names(Ar.geneList.int) %in% Ar.int))] <- 1
# check correct number of values set to 1
table(Ar.geneList.int)

@@

```{r rData, echo=FALSE}
# save RData for manuscript file
save(tt2, Ap.df, w1, w2, Ap.T_on, file = "ms.RData")
```
